# Attributes

{{< include _setup.qmd >}}

Almost every SEXP can carry attributes: named metadata stored as a
[pairlist](pairlists.qmd) hanging off the object (though you should treat
them as an opaque map and use the accessors below). Attributes are what
turn bare vectors into R's richer data structures — a matrix is a vector with a `dim` attribute, a
factor is an integer vector with `levels` and `class`, and a data frame is
a list with `names`, `row.names`, and `class`.

## Get and set

Attributes are stored as a tagged pairlist, but you should treat them as an
opaque map from symbol to value and use the accessors below.

```{r}
#| results: asis
#| echo: false
render_entries("attributes", "Get and set")
```

## Arrays

An array is an atomic vector with a `dim` attribute: an integer vector
giving the size of each dimension. Elements are stored in column-major
order (the first dimension varies fastest), so the flat data pointer from
`REAL()` etc. needs no rearrangement.

```{r}
#| results: asis
#| echo: false
render_entries("attributes", "Arrays")
```

## Matrices

A matrix is an array with exactly two dimensions. R provides a few
matrix-specific conveniences:

```{r}
#| results: asis
#| echo: false
render_entries("attributes", "Matrices")
```

## Factors

A factor is an integer vector of 1-based codes with a `levels` character
vector and class `"factor"`. `NA_INTEGER` codes represent missing values.
Because factors are `INTSXP` internally, `TYPEOF(x) == INTSXP` alone can't
distinguish them from plain integer vectors.

```{r}
#| results: asis
#| echo: false
render_entries("attributes", "Factors")
```

## Data frames

A data frame is a `VECSXP` of equal-length columns with `names`,
`row.names`, and class `"data.frame"`. The C API offers almost no support —
you build and manipulate them as plain lists with attributes.

The `row.names` attribute is special. It must be a character or integer
vector with one entry per row, but the default sequence `1:n` is stored in
a compact form: the length-2 integer vector `c(NA_integer_, -n)`.
`Rf_getAttrib(df, R_RowNamesSymbol)` expands this for you, returning a
freshly allocated length-`n` vector (an ALTREP compact sequence in recent
R). This makes it a rare case where `Rf_getAttrib()` allocates, so you must
`PROTECT` the result. When constructing a data frame in C you can set the
compact form yourself:

```c
SEXP rn = PROTECT(Rf_allocVector(INTSXP, 2));
INTEGER(rn)[0] = NA_INTEGER;
INTEGER(rn)[1] = -n;  // n rows
Rf_setAttrib(df, R_RowNamesSymbol, rn);
UNPROTECT(1);
```

## Time series

A `ts` object is a numeric vector or matrix with a `tsp` attribute
holding `c(start, end, frequency)`. This is mostly of historical
interest — most modern time series packages use their own classes — but
base R's `is.ts()` is available:

```{r}
#| results: asis
#| echo: false
render_entries("attributes", "Time series")
```
