# Environments

{{< include _setup.qmd >}}

* `R_GlobalEnv`: The "global" environment
* `R_EmptyEnv`: An empty environment at the root of the environment tree
* `R_BaseEnv`: The base environment; formerly R_NilValue
* `R_BaseNamespace`: The (fake) namespace for base
* `R_NamespaceRegistry`: Registry for registered namespaces
* `R_GetCurrentEnv`: Retrieve the current environment
* `R_NewEnv`: Create a new environment

Environments are commonly called `rho` in the sources.

## Predefined environments

```{r}
#| results: asis
#| echo: false
render_entries("environments", "Predefined environments")
```

## Creation

```{r}
#| results: asis
#| echo: false
render_entries("environments", "Creation")
```

## Get and set objects in environment

`symbol` should be a `SYMSXP`; `environment` should be an `ENVSXP`.

### Get values

Note that retrieving an variable from an environment may cause an allocation because it might be an active binding.

```{r}
#| results: asis
#| echo: false
render_entries("environments", "Get values")
```

### Set values

```{r}
#| results: asis
#| echo: false
render_entries("environments", "Set values")
```

To remove a binding, set its value to `R_UnboundSymbol`.

### Check for presence

```{r}
#| results: asis
#| echo: false
render_entries("environments", "Check for presence")
```

## Miscellaneous

```{r}
#| results: asis
#| echo: false
render_entries("environments", "Miscellaneous")
```

## Internals

```{r}
#| results: asis
#| echo: false
render_entries("environments", "Internals")
```

### Truelength

> This is almost unused. The only current use is for hash tables of 
> environments (VECSXPs), where length is the size of the table and 
> truelength is the number of primary slots in use, and for the reference 
> hash tables in serialization (VECSXPs), where truelength is the number of 
> slots in use.
> --- R-internals

```{r}
#| results: asis
#| echo: false
render_entries("environments", "Truelength")
```

### Hash

```{r}
#| results: asis
#| echo: false
render_entries("environments", "Hash")
```

// [R-Internals]: Hashing is principally designed for fast searching of
// environments, which  are from time to time added to but rarely deleted from,
// so items are not actually deleted but have their value set to R_UnboundValue.


Code from lobstr should help me understand how they work:

```c
int FrameSize(SEXP frame) {
  int count = 0;

  for(SEXP cur = frame; cur != R_NilValue; cur = CDR(cur)) {
    SEXP obj = CAR(cur)
    if (obj != R_UnboundValue)
      count++;
  }
  return count;
}

int HashTableSize(SEXP table) {
  int count = 0;
  int n = Rf_length(table);
  for (int i = 0; i < n; ++i)
    count += FrameSize(VECTOR_ELT(table, i));
  return count;
}

int envlength(SEXP x) {
  bool isHashed = HASHTAB(x) != R_NilValue;
  return isHashed ? HashTableSize(HASHTAB(x)) : FrameSize(FRAME(x));
}
```
