# Environments

{{< include _setup.qmd >}}

An environment (`ENVSXP`) is a set of bindings from symbols to values, plus
a pointer to a parent (enclosing) environment. Lookup walks the parent
chain until it finds a binding or reaches the empty environment. Small
environments store bindings as a pairlist (the frame); larger ones use a
hash table. In R's own source, environments are conventionally named `rho`.

## Predefined environments

```{r}
#| results: asis
#| echo: false
render_entries("environments", "Predefined environments")
```

## Creation

```{r}
#| results: asis
#| echo: false
render_entries("environments", "Creation")
```

## Get and set objects in environment

In these functions, `symbol` should be a `SYMSXP` and `environment` an
`ENVSXP`.

### Get values

Retrieving a variable from an environment can allocate, because the binding
might be an active binding or a promise that needs forcing.

```{r}
#| results: asis
#| echo: false
render_entries("environments", "Get values")
```

### Set values

To remove a binding, set its value to `R_UnboundValue`.

```{r}
#| results: asis
#| echo: false
render_entries("environments", "Set values")
```

### Check for presence

```{r}
#| results: asis
#| echo: false
render_entries("environments", "Check for presence")
```

## Miscellaneous

```{r}
#| results: asis
#| echo: false
render_entries("environments", "Miscellaneous")
```

## Inspecting bindings

R 4.6.0 added an **experimental** API for examining bindings in detail —
whether a binding is delayed (a promise), forced, missing, or active —
and for working with `...` without evaluating it. Both interface and
semantics may change in future R releases. Note that the `R_Dots*` and
`R_Dot*` accessors look up `...` only in the frame you pass; call
`R_findDotsEnv()` first if you want R's usual inherited lookup.

```{r}
#| results: asis
#| echo: false
render_entries("environments", "Inspecting bindings")
```
