---
aliases: [error-eval.html]
---

# Errors and evaluation

{{< include _setup.qmd >}}

## Errors, warnings, and printing

```{r}
#| results: asis
#| echo: false
render_entries("errors", "Errors, warnings, and printing")
```

Use `Rf_errorcall(R_NilValue, ...)` to suppress display of call.

Two macros eliminate compiler warnings about non-void functions that don't return.
(`Rf_error` will `longjmp` and so any code following will not be executed; however,
most compilers do not detect this when providing warnings)

## Evaluation

```{r}
#| results: asis
#| echo: false
render_entries("errors", "Evaluation")
```

### Protected evaluation

```{r}
#| results: asis
#| echo: false
render_entries("errors", "Protected evaluation")
```

Use these to evaluate R expressions in a stand-alone context ensuring that any errors produced during evaluation don't cause a longjmp to a separate context.

Note that these functions ignore all existing condition handlers so R functions like `supressWarnings()`, `tryCatch()`, will no longer work.

Both `R_tryEval()` and `R_tryEvalSilent`() call R_ToplevelExec under the hood:
