---
aliases: [error-eval.html]
---

# Errors and conditions

{{< include _setup.qmd >}}

R reports errors with a `longjmp`: `Rf_error()` never returns, abandoning
your C function on the spot. That has two consequences you must design
around. First, any cleanup your function owes — freeing `R_Calloc()`
memory, closing files, restoring global state — will not run unless you
arrange it in advance with `R_UnwindProtect()` or an external pointer
finalizer. Second, in C++ the longjmp skips destructors, so RAII objects
must live below an `R_UnwindProtect()` boundary that converts the longjmp
into a C++ exception (see `R_MakeUnwindCont()`).

For catching errors from R code you call, rather than signalling your own,
see the protected evaluation entry points in
[Evaluation](evaluation.qmd).

## Signalling

Follows [WRE §6.2, Error signaling](https://cran.r-project.org/doc/manuals/R-exts.html#Error-signaling-1)
closely.

`Rf_error()` and `Rf_warning()` take `printf`-style format strings. If
your message is a plain string that might contain `%`, pass it as
`Rf_error("%s", msg)` so it isn't interpreted as a format.

To signal an error with a custom class — so R code can catch it
selectively with `tryCatch(my_class = function(c) ...)` — there is no
dedicated entry point. Build a condition object (a named list with
`message` and `call` fields, classed `c("my_class", "error",
"condition")`) and signal it by evaluating a call to `stop()`:

```c
SEXP cls = PROTECT(Rf_allocVector(STRSXP, 3));
SET_STRING_ELT(cls, 0, Rf_mkChar("negative_value"));
SET_STRING_ELT(cls, 1, Rf_mkChar("error"));
SET_STRING_ELT(cls, 2, Rf_mkChar("condition"));

SEXP cond = PROTECT(Rf_mkNamed(VECSXP, (const char *[]){"message", "call", ""}));
SET_VECTOR_ELT(cond, 0, Rf_mkString("x must be non-negative"));
SET_VECTOR_ELT(cond, 1, R_NilValue);
Rf_classgets(cond, cls);

/* Never returns, so the UNPROTECT is never reached */
Rf_eval(Rf_lang2(Rf_install("stop"), cond), R_GlobalEnv);
```

The same pattern with `warning()` signals a classed warning.

```{r}
#| results: asis
#| echo: false
render_entries("errors", "Signalling")
```

## Condition handling and cleanup

Follows [WRE §6.13, Condition handling and cleanup code](https://cran.r-project.org/doc/manuals/R-exts.html#Condition-handling-and-cleanup-code-1)
closely.

These functions let C code do what `tryCatch()` and `on.exit()` do at the
R level: run a body function with a condition handler or cleanup action
installed. Reach for `R_UnwindProtect()` when you need guaranteed cleanup,
and the `R_tryCatch*()` family when you need to inspect or recover from a
condition object.

```{r}
#| results: asis
#| echo: false
render_entries("errors", "Condition handling and cleanup")
```

## Interrupts

Follows [WRE §6.14, Allowing interrupts](https://cran.r-project.org/doc/manuals/R-exts.html#Allowing-interrupts-1)
closely.

```{r}
#| results: asis
#| echo: false
render_entries("errors", "Interrupts")
```

## C stack checking

Follows [WRE §6.15, C stack checking](https://cran.r-project.org/doc/manuals/R-exts.html#C-stack-checking-1)
closely.

```{r}
#| results: asis
#| echo: false
render_entries("errors", "C stack checking")
```
