# Evaluation

{{< include _setup.qmd >}}

Evaluating R code from C is how you call back into R: apply a function,
access a method, or run user-supplied expressions. There is no public
"apply this closure to these arguments" entry point (`Rf_applyClosure()`
is internal); instead you build a call — a `LANGSXP` pairlist, see
[Pairlists, calls, and `...`](pairlists.qmd) — and evaluate it with
`Rf_eval()`.

Evaluation can do anything R code can do: allocate, signal errors, and
longjmp. Treat every `Rf_eval()` as a potential non-local return and
protect accordingly.

## Evaluation

Follows [WRE §5.11, Evaluating R expressions from C](https://cran.r-project.org/doc/manuals/R-exts.html#Evaluating-R-expressions-from-C)
closely.

A typical use looks up a function and calls it with C-constructed
arguments:

```c
SEXP call = PROTECT(Rf_lang2(Rf_install("sqrt"), x));
SEXP result = PROTECT(Rf_eval(call, R_GlobalEnv));
/* ... */
UNPROTECT(2);
```

For more than a couple of arguments, build the call with
`Rf_allocList()` + `SET_TYPEOF(..., LANGSXP)` or `Rf_cons()` rather than
deeply nesting `Rf_lang*()` calls. Evaluate in the environment that gives
the call the right scope: `R_GlobalEnv` for user-visible semantics, a
namespace environment for calling a package's internals, or
`R_BaseEnv`/`R_BaseNamespace` for base functions.

```{r}
#| results: asis
#| echo: false
render_entries("evaluation", "Evaluation")
```

## Protected evaluation

`Rf_eval()` longjmps on error, abandoning your C function mid-flight. If
you need to keep control after a failure — to report the error yourself,
retry, or clean up and continue — evaluate in a context that catches the
error instead. These functions ignore all existing condition handlers, so
R-level `tryCatch()` and `suppressWarnings()` have no effect on the
protected evaluation.

For richer control — running cleanup code on unwind, or installing
condition handlers — see [Errors and conditions](errors.qmd).

```{r}
#| results: asis
#| echo: false
render_entries("evaluation", "Protected evaluation")
```

## Parsing

Parsing turns text into an expression vector (`EXPRSXP`) without
evaluating it; evaluate the result element by element with `Rf_eval()`.
Always check the returned `ParseStatus` before evaluating.

```{r}
#| results: asis
#| echo: false
render_entries("evaluation", "Parsing")
```
