# Functions

{{< include _setup.qmd >}}

R has three function types: closures (`CLOSXP`, functions written in R),
builtins (`BUILTINSXP`), and specials (`SPECIALSXP`, both implemented in
C). They differ in how arguments are evaluated: builtins evaluate their
arguments before the call, specials don't. 

## Closures

Create closures with `R_mkClosure()` — the old idiom of allocating a
`CLOSXP` with `Rf_allocSExp()` and filling it with
`SET_FORMALS`/`SET_BODY`/`SET_CLOENV` is no longer part of the API.

```{r}
#| results: asis
#| echo: false
render_entries("functions", "Closures")
```

## Testing functions

```{r}
#| results: asis
#| echo: false
render_entries("functions", "Testing functions")
```

## Promises (`PROMSXP`)

A promise bundles an unevaluated expression with the environment it should
be evaluated in; it's how R implements lazy evaluation of function
arguments.

```{r}
#| results: asis
#| echo: false
render_entries("functions", "Promises (`PROMSXP`)")
```

## Srcrefs

```{r}
#| results: asis
#| echo: false
render_entries("functions", "Srcrefs")
```
