# Mathematical functions (Rmath)

{{< include _setup.qmd >}}

`Rmath.h` gives C code access to the routines behind R's mathematical and
statistical functions: distribution functions, special functions, and
numerical utilities, all with R's edge-case semantics (missing values,
`NaN`, infinities). The header has its own remapping scheme: define
`R_NO_REMAP_RMATH` and most names need an `Rf_` prefix — see the header
for which. We show the canonical (remapped) names below.

## Distribution functions

Follows [WRE §6.7.1, Distribution functions](https://cran.r-project.org/doc/manuals/R-exts.html#Distribution-functions-1)
closely.

Every distribution provides up to four entry points: `d` (density/mass),
`p` (CDF), `q` (quantile), and `r` (random generation). The table gives
the base name — prefix with `d`/`p`/`q`/`r` — and the
distribution-specific arguments, which come between the `x`/`p` argument
and the trailing `lower_tail`/`log_p` flags.

| Distribution | Base name | Parameters |
|---|---|---|
| beta | `beta` | `a`, `b` |
| non-central beta | `nbeta` | `a`, `b`, `ncp` |
| binomial | `binom` | `n`, `p` |
| Cauchy | `cauchy` | `location`, `scale` |
| chi-squared | `chisq` | `df` |
| non-central chi-squared | `nchisq` | `df`, `ncp` |
| exponential | `exp` | `scale` (not `rate`!) |
| F | `f` | `n1`, `n2` |
| non-central F | `nf` | `n1`, `n2`, `ncp` |
| gamma | `gamma` | `shape`, `scale` |
| geometric | `geom` | `p` |
| hypergeometric | `hyper` | `NR`, `NB`, `n` |
| logistic | `logis` | `location`, `scale` |
| lognormal | `lnorm` | `logmean`, `logsd` |
| negative binomial | `nbinom` | `size`, `prob` |
| normal | `norm` | `mu`, `sigma` |
| Poisson | `pois` | `lambda` |
| Student's t | `t` | `n` |
| non-central t | `nt` | `df`, `delta` |
| Studentized range | `tukey` | `rr`, `cc`, `df` (p/q only) |
| uniform | `unif` | `a`, `b` |
| Weibull | `weibull` | `shape`, `scale` |
| Wilcoxon rank sum | `wilcox` | `m`, `n` |
| Wilcoxon signed rank | `signrank` | `n` |

The non-central distributions have no `r` functions, and `tukey` has only
`p` and `q`. Note that the exponential and gamma distributions are
parametrized by `scale`, not `rate`.

```{r}
#| results: asis
#| echo: false
render_entries("math", "Distribution functions")
```

## Mathematical functions

Follows [WRE §6.7.2, Mathematical functions](https://cran.r-project.org/doc/manuals/R-exts.html#Mathematical-functions-1)
closely.

```{r}
#| results: asis
#| echo: false
render_entries("math", "Mathematical functions")
```

## Numerical utilities

Follows [WRE §6.7.3, Numerical utilities](https://cran.r-project.org/doc/manuals/R-exts.html#Numerical-Utilities-1)
closely.

Some of these (`log1p`, `expm1`, `cospi`, `sinpi`, `tanpi`) may be
provided by the platform's `math.h` instead of `Rmath.h`; under C++,
`math.h` is not included by `Rmath.h`, so declare them yourself or define
`__STDC_WANT_IEC_60559_FUNCS_EXT__` before the first inclusion.

```{r}
#| results: asis
#| echo: false
render_entries("math", "Numerical utilities")
```

## Mathematical constants

Follows [WRE §6.7.4, Mathematical constants](https://cran.r-project.org/doc/manuals/R-exts.html#Mathematical-constants-1)
closely.

| Name | Value | Name | Value |
|---|---|---|---|
| `M_E` | e | `M_SQRT2` | √2 |
| `M_LOG2E` | log2(e) | `M_SQRT1_2` | 1/√2 |
| `M_LOG10E` | log10(e) | `M_SQRT_3` | √3 |
| `M_LN2` | ln 2 | `M_SQRT_32` | √32 |
| `M_LN10` | ln 10 | `M_LOG10_2` | log10(2) |
| `M_PI` | π | `M_2PI` | 2π |
| `M_PI_2` | π/2 | `M_SQRT_PI` | √π |
| `M_PI_4` | π/4 | `M_1_SQRT_2PI` | 1/√(2π) |
| `M_1_PI` | 1/π | `M_SQRT_2dPI` | √(2/π) |
| `M_2_PI` | 2/π | `M_LN_SQRT_PI` | ln √π |
| `M_2_SQRTPI` | 2/√π | `M_LN_SQRT_2PI` | ln √(2π) |
| | | `M_LN_SQRT_PId2` | ln √(π/2) |

## Standalone Rmath

Follows [WRE §6.20, Using these functions in your own C code](https://cran.r-project.org/doc/manuals/R-exts.html#Using-these-functions-in-your-own-C-code)
closely.

Everything in this chapter is also available outside R: the same code
builds as a standalone library, `libRmath`, for use in programs that
don't link to R at all (define `MATHLIB_STANDALONE` when using it). The
library isn't built by default; see [The standalone Rmath
library](https://cran.r-project.org/doc/manuals/R-admin.html#The-standalone-Rmath-library)
in R Installation and Administration.
