# Memory allocation

{{< include _setup.qmd >}}

Sometimes you need scratch memory that isn't an R object — a work array, a
lookup table. R offers two kinds of allocation, differing in who cleans up:
**transient** memory that R reclaims automatically, and **user-controlled**
memory that you must free yourself, including on error paths. These
functions are declared in `R_ext/Memory.h` and `R_ext/RS.h` (both included
by `<R.h>`), and are not thread-safe.

## Transient allocation

Follows [WRE §6.1.1, Transient storage allocation](https://cran.r-project.org/doc/manuals/R-exts.html#Transient-storage-allocation-1)
closely.

`R_alloc()` memory comes from R's heap arena and is released automatically
when your `.Call` entry point returns — including when an error or
interrupt unwinds the stack. That makes it the default choice: it is
impossible to leak. If you call `R_alloc()` in a loop, bound the growth by
resetting the arena's high-water mark each iteration:

```c
void *vmax = vmaxget();
for (...) {
  double *work = (double *) R_alloc(n, sizeof(double));
  ...
  vmaxset(vmax);
}
```

The returned memory is only guaranteed to be aligned for `double`; use
`R_allocLD()` if you need `long double` alignment.

```{r}
#| results: asis
#| echo: false
render_entries("memory", "Transient allocation")
```

## User-controlled memory

Follows [WRE §6.1.2, User-controlled memory](https://cran.r-project.org/doc/manuals/R-exts.html#User_002dcontrolled-memory)
closely.

`R_Calloc()`/`R_Realloc()`/`R_Free()` are for memory that must outlive the
call or whose lifetime doesn't fit the arena model. The catch is the
longjmp contract ([Calling C from R](calling-c.qmd)): if R errors between
allocation and free, your `R_Free()` never runs and the memory leaks.
Options, in rough order of preference:

- Keep the allocation inside one function and wrap the body in
  `R_UnwindProtect()` with a cleanup handler that frees it.
- Free from an `on.exit()` action in the calling R function (base R's
  `pwilcox` does this).
- Attach the memory to an external pointer with a finalizer
  ([External pointers](external-pointers.qmd)) when it must persist
  across calls.

Also provided: `CallocCharBuf(n)` (a `char` buffer of `n + 1` for the NUL
terminator), and `Memcpy()`/`Memzero()` for copying and zeroing arrays.

```{r}
#| results: asis
#| echo: false
render_entries("memory", "User-controlled memory")
```

## When to use a RAWSXP instead

If the memory should be managed by R's garbage collector, visible to R, or
returned to R, don't use either of the above — allocate a `RAWSXP`
([Vectors](vectors.qmd)) and work in `RAW(x)`. You get GC-managed
lifetime, `PROTECT` semantics, and alignment suitable for any type, at the
cost of an R object header.
