# Numerical methods

{{< include _setup.qmd >}}

R exposes the C and Fortran code behind `optim()`, `integrate()`, and its
linear algebra routines, so package code can run the same algorithms
without going back through R. For pure scalar mathematics
(distributions, special functions), see
[Mathematical functions](math.qmd).

## Optimization

Follows [WRE §6.8, Optimization](https://cran.r-project.org/doc/manuals/R-exts.html#Optimization-1)
closely.

All five minimizers take the objective as a C callback, and some take a
gradient callback:

```c
typedef double optimfn(int n, double *par, void *ex);
typedef void optimgr(int n, double *par, double *gr, void *ex);
```

`ex` is an opaque pointer passed straight through to your callback — use
it to carry auxiliary data (including a `SEXP`, if you protect it
elsewhere). If you don't have an analytic gradient, you must
finite-difference yourself; no helper is provided.

```{r}
#| results: asis
#| echo: false
render_entries("numerical", "Optimization")
```

## Integration

Follows [WRE §6.9, Integration](https://cran.r-project.org/doc/manuals/R-exts.html#Integration-1)
closely.

The integrand is a *vectorized* callback — it receives `n` points in `x`
and must overwrite them with the function values:

```c
typedef void integr_fn(double *x, int n, void *ex);
```

In the current QUADPACK-based implementation `n` is always 15 or 21.

```{r}
#| results: asis
#| echo: false
render_entries("numerical", "Integration")
```

## Linear algebra

Follows [WRE §6.11, Linear algebra](https://cran.r-project.org/doc/manuals/R-exts.html#Linear-algebra-1)
closely.

For numerical linear algebra, call the BLAS/LAPACK routines that R itself
is built on: declarations are in `R_ext/BLAS.h` and `R_ext/LAPACK.h` (the
older `R_ext/Linpack.h` also exists). These are Fortran calling
conventions — all arguments are pointers, and character arguments need a
hidden length argument, which the `FCONE` macro (defined by these
headers) appends portably. This set of routines is not formally API but
changes rarely.

```{r}
#| results: asis
#| echo: false
render_entries("numerical", "Linear algebra")
```

## Miscellaneous

```{r}
#| results: asis
#| echo: false
render_entries("numerical", "Miscellaneous")
```
