---
aliases: [gc-rc.html]
---

# Protection and the garbage collector

{{< include _setup.qmd >}}

R's garbage collector reclaims `SEXP`s that are no longer reachable. That
creates a hazard unique to writing C against R: a `SEXP` you just allocated
can be freed out from under you at the *next allocation*, unless you tell R
you're still using it. Telling R is called **protection**, and getting it
right is the single most important discipline in R C programming.

## When the garbage collector runs

GC can only run when R allocates — `Rf_allocVector()`, `Rf_cons()`,
`Rf_install()`, `Rf_mkString()`, and anything else that creates an object.
Pure accessors (`REAL()`, `TYPEOF()`, `CAR()`) never trigger it. The
corollary: an unprotected `SEXP` is safe only until the next allocation,
which makes protection bugs easy to miss — they surface only when GC
happens to run at the wrong moment.

```{r}
#| results: asis
#| echo: false
render_entries("protection", "GC")
```

## What needs protecting

Protect every `SEXP` you allocate, unless it is immediately stored inside
an already-protected object. You do **not** need to protect:

- **Function arguments** — R is holding them for the duration of the call.
- **Symbols** — interned symbols are permanently reachable.
- **Values reachable from a protected object** — e.g. once you've
  `SET_VECTOR_ELT(protected_list, i, x)`, `x` is safe.
- **`R_NilValue`** and other R-owned constants.

When in doubt, protect: an extra `PROTECT` is cheap, a missing one is a
crash.

## PROTECT and UNPROTECT

`PROTECT()` pushes the object onto a protection stack (an internal array of
`SEXP`s); `UNPROTECT(n)` pops the `n` most recent entries. Protects and
unprotects must balance exactly — R warns about a "stack imbalance" when
they don't. Because the stack is only reset when your `.Call` entry point
returns, protecting inside a loop without unprotecting can overflow it;
use `PROTECT_WITH_INDEX()`/`REPROTECT()` to reuse one stack slot across
iterations.

```{r}
#| results: asis
#| echo: false
render_entries("protection", "Protect")
```

## Preserving objects across calls

`PROTECT` only lasts until your `.Call` returns. To keep an object alive
across multiple calls from R (typically stashed in a static variable or an
external pointer), add it to R's precious list. The precious list is a
pairlist walked by the GC; because `R_ReleaseObject()` searches it
linearly, preserving and releasing many objects in the same order can
degrade to quadratic time — see
[this dplyr issue](https://github.com/tidyverse/dplyr/issues/1396) for a
real-world manifestation.

```{r}
#| results: asis
#| echo: false
render_entries("protection", "R_PreserveObject, R_ReleaseObject")
```

## Modifying inputs and reference counting

You must not modify a function argument in place: it may be shared with
the caller (see [SEXPs](sexps.qmd)). Duplicate first — for lists,
`Rf_shallow_duplicate()` copies only the top level, while `Rf_duplicate()`
copies every element. `MAYBE_SHARED()` and friends let you check the
reference-count flags before modifying an object you *did* create, which
is the optimisation R's own internals use.

```{r}
#| results: asis
#| echo: false
render_entries("protection", "Reference counting")
```

```{r}
#| results: asis
#| echo: false
render_entries("protection", "Efficient copying")
```

## Finding protection bugs

Protection bugs typically manifest as intermittent segfaults or corrupted
objects. Two tools make them reproducible:

- `gctorture(TRUE)` (or `gctorture2()` for finer control) forces a GC at
  every allocation, so a missing `PROTECT` fails immediately instead of
  randomly. Combine with `gctorture(TRUE); <your test>` under
  `R CMD check`.
- [rchk](https://github.com/kalibera/rchk) statically analyses your
  compiled package for protect/unprotect imbalances and unprotected
  uses. It has false positives, but every report is worth reading.

## Header flags

The macros below read and write the flags in an object's header. You
rarely need them directly — prefer the higher-level accessors noted in
the entry.

```{r}
#| results: asis
#| echo: false
render_entries("protection", "Internals")
```
