---
aliases: [save-load.html]
---

# Serialisation

{{< include _setup.qmd >}}

R's serialisation format converts an R object to a byte stream (and back)
that can be written to a file, a connection, or a database. From C you can
drive the same machinery `serialize()`/`unserialize()` use at the R level,
either over a `FILE *` or over a custom byte source/sink you provide.

## XDR encoding

XDR is the default binary serialisation format: big-endian and
platform-independent, so streams are portable across architectures. These
helpers encode and decode individual values when you implement a custom
stream.

```{r}
#| results: asis
#| echo: false
render_entries("serialisation", "XDR encoding")
```

## Persistence streams

A persistence stream pairs a byte source/sink with optional persistence
hooks. The hook function (`phook`, with its data `pdata`) is called for
objects that can't be serialised by value — external pointers and weak
references — letting you substitute a placeholder on output and
reconstruct the object on input; pass `NULL` to refuse such objects with
an error. The `version` argument selects the serialisation format version
(2, or 3 which supports ALTREP; version 3 output requires R >= 3.6.0 to
read).

Note that WRE describes this framework as **highly experimental** for
package use: it may change or be dropped in a future R release. Don't
depend on the fields of the stream structs directly — initialise them with
the functions below.

```c
typedef void *R_pstream_data_t;

typedef enum {
  R_pstream_any_format,
  R_pstream_ascii_format,
  R_pstream_binary_format,
  R_pstream_xdr_format,
  R_pstream_asciihex_format
} R_pstream_format_t;

typedef struct R_outpstream_st *R_outpstream_t;
struct R_outpstream_st {
  R_pstream_data_t data;
  R_pstream_format_t type;
  int version;
  void (*OutChar)(R_outpstream_t, int);
  void (*OutBytes)(R_outpstream_t, void *, int);
  SEXP (*OutPersistHookFunc)(SEXP, SEXP);
  SEXP OutPersistHookData;
};

typedef struct R_inpstream_st *R_inpstream_t;
struct R_inpstream_st {
  R_pstream_data_t data;
  R_pstream_format_t type;
  int (*InChar)(R_inpstream_t);
  void (*InBytes)(R_inpstream_t, void *, int);
  SEXP (*InPersistHookFunc)(SEXP, SEXP);
  SEXP InPersistHookData;
};
```

The connection interface is not available to packages. To allow limited
use of connection pointers, defining `NEED_CONNECTION_PSTREAMS` before
including `Rinternals.h` defines the opaque pointer type:

```c
#ifdef NEED_CONNECTION_PSTREAMS
#ifndef HAVE_RCONNECTION_TYPEDEF
typedef struct Rconn  *Rconnection;
#define HAVE_RCONNECTION_TYPEDEF
#endif
#endif
```

```{r}
#| results: asis
#| echo: false
render_entries("serialisation", "Persistence streams")
```

## Serialising objects

With a stream initialised, these two functions do the actual work:

```{r}
#| results: asis
#| echo: false
render_entries("serialisation", "Serialising objects")
```
