# Strings

{{< include _setup.qmd >}}

A character vector (`STRSXP`) is an array of pointers to `CHARSXP`s, and
every `CHARSXP` lives in a global string pool. This means each distinct
string is stored only once, however many character vectors reference it —
and it means you should never modify the contents of a `CHARSXP`.

## Encodings

Each `CHARSXP` carries an encoding flag. R assumes strings are in the
native encoding unless marked otherwise; UTF-8 is the safe default for new
code. See also [Querying CHARSXP
encoding](https://cran.r-project.org/doc/manuals/R-exts.html#Querying-CHARSXP-encoding-1)
in Writing R Extensions.

```{r}
#| results: asis
#| echo: false
render_entries("strings", "Encodings")
```

## Creating from C strings

These calls create `CHARSXP`s (or length-1 `STRSXP`s) from C strings.
Typically they don't need protection because the result is immediately
assigned into a `STRSXP`.

They will crash R if passed `NULL`; check for it yourself and substitute
`""` or `NA_STRING` as appropriate.

Most modern C libraries produce UTF-8, so you should typically use
`Rf_mkCharCE()` or `Rf_mkCharLenCE()` with `CE_UTF8`, and avoid the
native-encoding creation functions, including `Rf_mkString()`.

To re-encode strings from another encoding, use R's wrapper around iconv
in `R_ext/Riconv.h`, which provides cross-platform `Riconv_open()` and
`Riconv()`. It's usually best to convert to UTF-8. See
[Re-encoding](https://cran.r-project.org/doc/manuals/R-exts.html#Re_002dencoding-1)
in Writing R Extensions.

```{r}
#| results: asis
#| echo: false
render_entries("strings", "Creating from C strings")
```

## Convert to C string

To access the C string stored in a `CHARSXP`, use `CHAR()`. (To match the
other vector accessors like `INTEGER()`, this is typically called via the
`Rf_`-less name.)

The `Rf_translateChar*` functions return a `const char*` in the specified
encoding; most modern C APIs use UTF-8, so you almost always want
`Rf_translateCharUTF8()`.

If re-encoding is necessary, the returned `char*` is allocated with
`R_alloc()` and freed automatically at the end of the `.Call`; if you need
it across calls, make a copy. If you make many translating calls in a loop,
free the transient memory explicitly with `vmaxget()`/`vmaxset()`:

```c
const void *vmax = vmaxget();
... // one or more calls to Rf_translateCharUTF8(), etc.
vmaxset(vmax);
```

```{r}
#| results: asis
#| echo: false
render_entries("strings", "Convert to C string")
```

## Special values

```{r}
#| results: asis
#| echo: false
render_entries("strings", "Special values")
```

## Helper functions

```{r}
#| results: asis
#| echo: false
render_entries("strings", "Helper functions")
```

### Paths

```{r}
#| results: asis
#| echo: false
render_entries("strings", "Paths")
```

### Types

You can convert `SEXPTYPE`s to and from C strings:

```{r}
#| results: asis
#| echo: false
render_entries("strings", "Types")
```
