# Strings

{{< include _setup.qmd >}}

A character vector (`STRSXP`) is effectively an array of pointers to `CHARSXP`s, which are stored in a global string pool. This design allows individual `CHARSXP`'s to be shared between multiple character vectors, reducing memory usage. 

## Encodings

Each `CHARSXP` has an encoding associated with it:

```{r}
#| results: asis
#| echo: false
render_entries("strings", "Encodings")
```

## Creating from C strings

```{r}
#| results: asis
#| echo: false
render_entries("strings", "Creating from C strings")
```

(Typically these calls don't need to be protected as they are immediately assigned into a `STRSXP`.)

If you need to re-encode strings from another encoding, use R's wrapper around iconv found in `R_ext/Riconv.h`. It provides cross-platform `Riconv_open()` and `Riconv`. It's usually best to convert to UTF-8.

Note that these calls will crash R if passed a `NULL`. It is your responsibility to check for that and return the correct value (typically `""` or `NA_STRING`).

Note that most modern C libraries encode strings as UTF-8. This means you should typically use `Rf_mkCharCE()` or `Rf_mkCharLenCE()`, and avoid the other string creation methods including `Rf_mkString()`

## Convert to C string

To access the underlying C-string stored in a CHARSXP, use `CHAR()`:

```{r}
#| results: asis
#| echo: false
render_entries("strings", "Convert to C string")
```

__NB__: To match the other vector accessor functions (`LOGICAL()`, `INTEGER()`, ...), this is typically called as `CHAR()`.

Three functions create `const char*` with the specified encoding.

Most modern C APIs use UTF-8, so you almost always want `Rf_translateCharUTF8()`.

__NB__: if re-encoding is necessary, the `char*` will be allocated by `R_alloc()` and will be automatically freed at the end of the call to `.C`, `.Call`, or `.External`. If you want to save across calls to C, make a copy.

If the memory is needed only for a short period of time or if you are making a large number of repeated calls that allocate memory, you may want to manage the deallocation explicitly. This is alluded to in the R source code with this comment (`src/main/sysutils.c`):
```c
/* This may return a R_alloc-ed result, so the caller has to manage the
   R_alloc stack */
const char *translateChar(SEXP x)
{
...
```

The code would look like this:
```c
const void *vmax = vmaxget();
... // one or more calls to Rf_translateChar(), Rf_translateCharUTF8(), etc.
vmaxset(vmax);
```


## Special values

```{r}
#| results: asis
#| echo: false
render_entries("strings", "Special values")
```

## Helper functions

```{r}
#| results: asis
#| echo: false
render_entries("strings", "Helper functions")
```

### Paths

```{r}
#| results: asis
#| echo: false
render_entries("strings", "Paths")
```

### Types

You can convert `SEXPTYPE`s to and from C strings with:

```{r}
#| results: asis
#| echo: false
render_entries("strings", "Types")
```
