# Symbols

{{< include _setup.qmd >}}

A symbol (`SYMSXP`) is a name, like `x` or `+`. Symbols are interned:
there is exactly one `SYMSXP` per name, stored in a global table, so two
symbols are equal iff they are the same pointer. Because they're never
garbage collected, symbols don't need protection, and if you use one
repeatedly it's worth caching the result of `Rf_install()`.

## Test

```{r}
#| results: asis
#| echo: false
render_entries("symbols", "Test")
```

## Create

`Rf_install()` returns the existing symbol for a name, or creates it if it
doesn't exist yet.

```{r}
#| results: asis
#| echo: false
render_entries("symbols", "Create")
```

## Accessors

```{r}
#| results: asis
#| echo: false
render_entries("symbols", "Accessors")
```

## Missing symbol

`R_MissingArg` represents the missing/empty symbol, i.e. the second
argument in the call `f(x, )`.

```{r}
#| results: asis
#| echo: false
render_entries("symbols", "Missing symbol")
```

## Predefined symbols

A number of symbols are so commonly used that they're predefined and
exported as global constants — use these instead of calling `Rf_install()`
yourself.

```{r}
#| results: asis
#| echo: false
render_entries("symbols", "Predefined symbols")
```
