---
aliases: [misc.html]
---

# Utilities

{{< include _setup.qmd >}}

Assorted entry points that don't fit elsewhere: options, parsing, and
file paths. Most are declared in `R_ext/Utils.h` (included by `<R.h>`).
For sorting, ordering, matching, and duplication, see
[Vectors](vectors.qmd); for string searching, see
[Strings](strings.qmd).

## Testing

```{r}
#| results: asis
#| echo: false
render_entries("utilities", "Testing")
```

## Options

```{r}
#| results: asis
#| echo: false
render_entries("utilities", "Options")
```

## Numeric parsing

`R_atof()`/`R_strtod()` exist because the C library equivalents are
locale-dependent: R's versions always use `.` as the decimal point and
recognise `"NA"`.

```{r}
#| results: asis
#| echo: false
render_entries("utilities", "Numeric parsing")
```

## Files and system

```{r}
#| results: asis
#| echo: false
render_entries("utilities", "Files and system")
```

## Platform information

Follows [WRE §6.17, Platform and version information](https://cran.r-project.org/doc/manuals/R-exts.html#Platform-and-version-information-1)
closely.

A few macros help write platform-aware code: `USING_R` confirms the code
is being compiled for R, and `Rconfig.h` (included by `<R.h>`) defines
platform macros such as `WORDS_BIGENDIAN` — but note these describe the
compiler that built R, not necessarily yours. For compile-time R version
checks (`R_VERSION`, `R_Version()`), see [R version](r-version.qmd). If
you use `alloca`, define it portably via `Rconfig.h`'s `HAVE_ALLOCA_H`
as shown in WRE §6.17.
