# Vectors

{{< include _setup.qmd >}}

R has six atomic vector types, plus two vector types whose elements are
themselves SEXPs:

* Logical (`LGLSXP`): `int`, using `0`/`1` for false/true and `NA_LOGICAL`
  for missing.
* Integer (`INTSXP`): `int`, with `NA_INTEGER` for missing.
* Double (`REALSXP`): `double`.
* Complex (`CPLXSXP`): `Rcomplex`.
* Raw (`RAWSXP`): `Rbyte`.
* String (`STRSXP`): elements are `CHARSXP`s; see [Strings](strings.qmd).
* List (`VECSXP`): elements are arbitrary SEXPs. Beware: lists are
  `VECSXP`s, not `LISTSXP`s — early R lists were Lisp-like linked lists,
  which survive as [pairlists](pairlists.qmd).
* Expression (`EXPRSXP`): like a list, but intended to hold language objects.

Two supporting C types:

```c
typedef unsigned char Rbyte;

typedef struct {
  double r;
  double i;
} Rcomplex;
```

## Length

After the SEXPTYPE, the most important property of a vector is its length.
There are two length types:

```c
typedef int R_len_t;          // limited to 2^31 - 1
typedef ptrdiff_t R_xlen_t;   // 64-bit on all supported platforms
```

New code should always use `R_xlen_t` for vector lengths and indices. It
costs nothing on 64-bit platforms, and the alternative fails badly: an
`int` length silently overflows if a user passes a vector longer than
$2^{31}-1$. Reserve `int` for things that are inherently small, like the
number of arguments to a call. You'll still need occasional casts where
older parts of the API take or return `int`.

Vectors also carry a hidden "truelength" field used for over-allocation:
since R 3.4.0, assigning past the end of a vector grows the allocation by
~5% and records the capacity in truelength, so repeated appends can extend
the vector in place. The experimental resizable-vector API
(`R_allocResizableVector()`/`R_resizeVector()`) builds on the same
mechanism.

```{r}
#| results: asis
#| echo: false
render_entries("vectors", "Length")
```

## Create

The most common way to create a new vector is `Rf_allocVector()`. As with
all SEXP creation functions, you must `PROTECT()` the result unless you
immediately assign it into an already-protected object. Alternatively, you
can coerce an existing vector with `Rf_coerceVector()`.

To create a vector of length 1 from a C value, use the scalar helpers —
they're shorter and clearer than allocating and assigning separately:

```c
SEXP x = PROTECT(Rf_ScalarReal(3.14));
// equivalent to:
// SEXP x = PROTECT(Rf_allocVector(REALSXP, 1));
// REAL(x)[0] = 3.14;
UNPROTECT(1);
```

`Rf_mkNamed()` creates a named vector in one call. This template creates a
list holding two objects `x` and `y` named `xname` and `yname`:

```c
// array of names; note the terminating empty string
const char *names[] = {"xname", "yname", ""};
SEXP list = PROTECT(Rf_mkNamed(VECSXP, names));  // creates a list of length 2
SET_VECTOR_ELT(list, 0, x); // x and y are arbitrary SEXPs
SET_VECTOR_ELT(list, 1, y);
UNPROTECT(1);
```

Finally, `Rf_allocVector3()` lets you supply a custom allocator:

```c
typedef void *(*custom_alloc_t)(R_allocator_t *allocator, size_t);
typedef void  (*custom_free_t)(R_allocator_t *allocator, void *);
typedef struct R_allocator {
    custom_alloc_t mem_alloc; /* malloc equivalent */
    custom_free_t  mem_free;  /* free equivalent */
    void *res;                /* reserved (maybe for copy) - must be NULL */
    void *data;               /* custom data for the allocator implementation */
} R_allocator_t;
```

```{r}
#| results: asis
#| echo: false
render_entries("vectors", "Create")
```

## Get and set values

The atomic types wrap a C array, so you get and set values through a helper
that returns a pointer to the underlying data. These helpers are all
uppercase and don't start with `Rf_`.

When working with longer vectors, cache the pointer instead of calling the
accessor on every iteration. Instead of:

```c
for (R_xlen_t i = 0; i < n; ++i) {
  INTEGER(x)[i] = INTEGER(x)[i] * 2;
}
```

do:

```c
int* px = INTEGER(x);
for (R_xlen_t i = 0; i < n; ++i) {
  px[i] = px[i] * 2;
}
```

Strings and lists don't map to simple C arrays, so they have paired
get/set functions (`STRING_ELT()`/`SET_STRING_ELT()`,
`VECTOR_ELT()`/`SET_VECTOR_ELT()`). (`STRING_PTR()` exists but is used only
in a handful of places in R itself; `VECTOR_PTR()` is a deprecated
interface that now throws an error.)

```{r}
#| results: asis
#| echo: false
render_entries("vectors", "Get and set values")
```

### Scalars

A few helpers extract the first value of a vector, coercing as necessary:

```{r}
#| results: asis
#| echo: false
render_entries("vectors", "Scalars")
```

### Special values

Integer, logical, and character vectors use sentinel values for missing
data (`NA_INTEGER`, `NA_LOGICAL`, `NA_STRING`). `NA_INTEGER` is `INT_MIN`,
the smallest representable `int` — which means R integer vectors cannot
store that value, and negating or taking the absolute value of `NA_INTEGER`
overflows.

Missing values are more complicated for `REALSXP` because doubles already
have a missing-value protocol defined by the floating-point standard
([IEEE 754](http://en.wikipedia.org/wiki/IEEE_floating_point)). An `NA` is
a `NaN` with a special bit pattern (the lowest word is 1954, the year Ross
Ihaka was born), and there are distinct values for positive and negative
infinity. Test with the `ISNA()`, `ISNAN()`, and `!R_FINITE()` macros; set
with the constants `NA_REAL`, `R_NaN`, `R_PosInf`, and `R_NegInf`.

```{r}
#| results: asis
#| echo: false
render_entries("vectors", "Special values")
```

## Test

A number of helpers test whether a SEXP has a given type. There's no
`Rf_isRaw()` or `Rf_isList()` for `RAWSXP`/`VECSXP`; use
`TYPEOF(x) == RAWSXP` or `TYPEOF(x) == VECSXP`. (`Rf_isList()` tests for a
pairlist.)

These tests are not always consistent with their R equivalents. For
example, `Rf_isVectorAtomic(R_NilValue)` is false but `is.atomic(NULL)` is
true; `Rf_isNewList(R_NilValue)` is true but `is.list(NULL)` is false.
Because of this confusion, prefer writing your own wrappers around
`TYPEOF(x)`.

```{r}
#| results: asis
#| echo: false
render_entries("vectors", "Test")
```

## Sorting and ordering

Follows [WRE §6.10, Utility functions](https://cran.r-project.org/doc/manuals/R-exts.html#Utility-functions-1)
closely.

Two families: the `R_*sort` functions sort C arrays in place (the
`R_qsort*` variants use 1-based `i:j` ranges and don't handle `NA`;
`R_isort`/`R_rsort`/`R_csort` sort `NA`s last), while `R_orderVector()`
computes an ordering permutation of R vectors, like `order()` — note it
returns 0-based indices. None of the in-place sorts are stable.

```{r}
#| results: asis
#| echo: false
render_entries("vectors", "Sorting and ordering")
```

## Matching and duplication

Vector-level equivalents of `match()`, `duplicated()`, and
`anyDuplicated()`.

```{r}
#| results: asis
#| echo: false
render_entries("vectors", "Matching and duplication")
```

## Miscellaneous helpers

A couple of macros test whether an object is a "scalar" (a vector of
length 1):

```{r}
#| results: asis
#| echo: false
render_entries("vectors", "Miscellaneous helpers")
```
