# Vectors

{{< include _setup.qmd >}}

There are seven vector types in R:

* Logical (`LGLSXP`), contains `Rboolean`.

* Integer (`INTSXP`), contains `int`.

* Double  (`REALSXP`), contains `double`.

* Complex (`CPLXSXP`), contains `Rcomplex`.

* String  (`STRSXP`), contains `CHARSXP`.

* Lists   (`VECSXP`), contains any other sexp. __Beware:__ Lists are `VECSXP`s
  not `LISTSXP`s. This is because early implementations of lists were Lisp-like 
  linked lists, which are now called as "[pairlists](pairlists.qmd)".

* Raw     (`RAWSXP`), contains `Rbyte`.

* Expression (`EXPRSXP`), contains `LANGSXP`, `SYMSXP` or a vector 
  (except for a list).

Three of those vectors are built on top of simple types defined by R:

```c
typedef unsigned char Rbyte;

typedef enum { 
  FALSE = 0, 
  TRUE 
} Rboolean;

typedef struct {
  double r;
  double i;
} Rcomplex;
```

## Length

```{r}
#| results: asis
#| echo: false
render_entries("vectors", "Length")
```


After the SEXPTYPE, the most important property of a vector is it's length. Historically, R vectors were limited to length $2 ^ {31} -  1$. Still most vectors are shorter than this, so you can use an `int` based interface:

```c
typedef int R_len_t;
```

As of R 3.0.0, R vectors can have length up to $2 ^ {64} - 1$. If you want your code to be as general as possible, you should instead use the `R_xlen_t`based interface:

```c
// ptrdiff_t is the type of the result of subtracting two pointers, and
// is usually 8 bytes (like a double)
typedef ptrdiff_t R_xlen_t;
```

These functions also have uppercase variants - in base R code these are implemented as macros for efficiency.

## Create

```{r}
#| results: asis
#| echo: false
render_entries("vectors", "Create")
```


The most common way to create a new vector is with `Rf_allocVector()`.
If you want to create a vector of length 1 from the corresponding C type, use the scalar helpers.

(Note, as with all SEXP creation functions you must `Rf_protect()` the result of any of these calls unless you're immediately assigning it into an already protected object)

Alternatively you can coerce from an existing vector with `Rf_coerceVector()`.

To create a named vector, use `Rf_mkNamed()`.

For example, this template creates a list holding two objects `x` and `y` named `xname` and `yname`:

```c
  // array of names; note the null string
  const char *names[] = {"xname", "yname", ""};
  SEXP list = PROTECT(Rf_mkNamed(VECSXP, names));  // creates a list of length 2
  SET_VECTOR_ELT(list, 0, x); // x and y are arbitrary SEXPs
  SET_VECTOR_ELT(list, 1, y);
  UNPROTECT(1);
```

Finally, you can use a custom allocator:
```c
// Create a vector with a custom memory allocator
typedef void *(*custom_alloc_t)(R_allocator_t *allocator, size_t);
typedef void  (*custom_free_t)(R_allocator_t *allocator, void *);
typedef struct R_allocator {
    custom_alloc_t mem_alloc; /* malloc equivalent */
    custom_free_t  mem_free;  /* free equivalent */
    void *res;                /* reserved (maybe for copy) - must be NULL */
    void *data;               /* custom data for the allocator implementation */
} R_allocator_t;
```

## Get and set values

```{r}
#| results: asis
#| echo: false
render_entries("vectors", "Get and set values")
```

The simple types are wrappers around an array of C values, so you get and set by using a helper that returns a pointer:

__NB__: these functions are all upper case and don't start with `Rf_`.

When working with longer vectors, there's typically a performance advantage to saving the index and reusing. For example, instead of:

```c
for (int i = 0; i < n; ++i) {
  INTEGER(x)[i] = INTEGER(x)[i] * 2;
}
```

Do:

```c
int* px = INTEGER(x);
for (int i = 0; i < n; ++i) {
  px[i] = px[i] * 2;
}
```

Strings and lists don't map to simple C structs, so instead have a pair of functions to get and set values.

(There is `STRING_PTR()` which is used in a handful of places in the R source; `VECTOR_PTR()` is a deprecated interface that now throws an error.)

### Scalars

```{r}
#| results: asis
#| echo: false
render_entries("vectors", "Scalars")
```

There are a few helpers that extract the first value, coercing the vector as necessary:

### Special values

```{r}
#| results: asis
#| echo: false
render_entries("vectors", "Special values")
```

Integer, logical, and character vectors have special sentinels for missing values. 

Missing values are somewhat more complicated for `REALSXP` because there is an existing protocol for missing values defined by the floating point standard ([IEEE 754](http://en.wikipedia.org/wiki/IEEE_floating_point)). In doubles, an `NA` is `NaN` with a special bit pattern (the lowest word is 1954, the year Ross Ihaka was born), and there are other special values for positive and negative infinity. Use `ISNA()`, `ISNAN()`, and `!R_FINITE()` macros to check for missing, NaN, or non-finite values. Use the constants `NA_REAL`, `R_NaN`, `R_PosInf`, and `R_NegInf` to set those values.

## Test

```{r}
#| results: asis
#| echo: false
render_entries("vectors", "Test")
```

A number of helpers let you test if an SEXP is of the given type:

__NB__: there's no function to test for `RAWSXP` or `VECSXP`; you use must use `TYPEOF(x) == RAWSXP`, or `TYPEOF(x) == VECSXP`. `isList()` tests if the object is a pairlist.

A handful of functions test for frequently used combinations of variable types:

__NB__: these are not always consistent with their R equivalents. For example, `Rf_isVectorAtomic(R_NilValue)` is false, but `is.atomic(NULL)` is true; `Rf_isNewList(R_NilValue)` is true; but `is.list(NULL)` is false. Because of this confusion, I recommend writing your own wrapper around `TYPEOF(x)`.

__NB__: Be careful when checking whether whether an R object is an integer vector. Internally, `factor`s are also `INTSXP`s so `TYPEOF(x) == INTSXP` would accept both vanilla integer vectors and factors. (Prefer using `Rf_isInteger()` and `Rf_isFactor()` for these cases.)

__NB__: It is somewhat strange that `Rf_isNewList()` returns `TRUE` for `R_NilValue`, given that `R_NilValue` is just an empty pairlist (ie, `NULL` and `pairlist()` are identical).

# Variants

The R API provides some support for the various data structures built on top of vectors.

## Arrays

```{r}
#| results: asis
#| echo: false
render_entries("vectors", "Arrays")
```

Arrays are vectors with a dim attribute:

### Matrices

```{r}
#| results: asis
#| echo: false
render_entries("vectors", "Matrices")
```

Matrices are a special case of arrays; those with 2 dimensions:

### Factors

```{r}
#| results: asis
#| echo: false
render_entries("vectors", "Factors")
```

### Data frame

```{r}
#| results: asis
#| echo: false
render_entries("vectors", "Data frame")
```

## Miscellaneous helpers

```{r}
#| results: asis
#| echo: false
render_entries("vectors", "Miscellaneous helpers")
```

A couple of macros aid in testing if an object is a "scalar" (a vector of length 1):
